Notes on Big MRA

### 2018-05-14 TEX/notes_will/notes_moments/backups/test2_save10.tex
updates on section 2

\newcommand{\F}{\mathcal{F}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\ep}{\varepsilon}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\GCD}{\mathbf{GCD}}

@@

We will assume that the first and last copies of $x$ is at least $L$ places removed from the boundaries of the interval $[1,n + L]$ (if not we can just zero-pad the interval). We will also denote by $I_1,\dots,I_J$ the $J$ subintervals containing the signal; and take $I_j = [a_j,b_j]$. When we take the limit $n \to \infty$, we'll let $J = J_n$ grow with $n$, as is natural. We will see that we require $J_n = \Omega(n)$ in order for the limits to not vanish.

@@

We fix a value $\Delta$ between $1$ and $L-1$. We will compute the second moment:
%
\begin{align}
%
    M_2(\Delta) = 
        \lim_{n \to \infty} \frac{1}{n} \sum_{i=1}^{n-\Delta} Y_i Y_{i+\Delta}.
%
\end{align}

First, if there is no noise, and under the wide spacing assumption, we can break the sum into $J_n$ different sums, one for each copy of $x$ embedded in the signal:
%
\begin{align}
%
    M_2(\Delta) 
        = \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=a_j}^{b_j-\Delta} Y_i Y_{i+\Delta}
    = \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=1}^{L-\Delta} X_i X_{i+\Delta}
%
\end{align}

@@

Here, $Y_i = \ep_i$. We fix two indices $k$ and $l$, and we suppose $k < l$. We will let $\Delta = l-k > 0$. We want to compute the limit $\frac{1}{n} \sum_{i=1}^n Y_{i + k - 1} Y_{i + l - 1} = \frac{1}{n} \sum_{i=1}^n Y_{i + k - 1} Y_{i + k + \Delta - 1}$. Break up the sum into $\Delta$ terms as follows:
%
\begin{align}
%
    &\frac{1}{n} \sum_{i=1}^n Y[i + k - 1]  Y[i + k+\Delta - 1]   
    \nonumber \\
    =& \frac{1}{\Delta} \sum_{m=0}^{\Delta-1} \frac{1}{n/\Delta}\sum_{j=1}^{n/\Delta} 
                      Y[j\Delta + m] Y[(j + 1)\Delta + m].
%
\end{align}
%
Each term
%
\begin{math}
    \frac{1}{n/\Delta}\sum_{j=1}^{n/\Delta} 
                      Y[j\Delta + m] Y[(j + 1)\Delta + m]
\end{math}
%
is an average of $n / \Delta$ iid terms with expectation zero, and so converges to 0 a.s.\ as $n \to \infty$. (Of course, it's not exactly $n / \Delta$ terms; there will be finitely many terms unaccounted for this way, but these are negligible).

If $k = l$, then it's even easier:
%
\begin{align}
%
    \frac{1}{n} \sum_{i=1}^n Y[i + k - 1]^2 \operatorname*{\longrightarrow}^{a.s.} \sigma^2.
%
\end{align}

So the second moment matrix of pure noise converges a.s.\ as $n \to \infty$ to $\sigma^2 I_L$.

>>>>

\newcommand{\F}{\mathcal{F}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\ep}{\varepsilon}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\GCD}{\mathbf{GCD}}
\newcommand{\XX}{\mathcal{X}}

@@

We will assume that the first and last copies of $x$ is at least $L$ places removed from the boundaries of the interval $[1,n + L]$ (if not we can just zero-pad the interval). We will also denote by $I_1,\dots,I_J$ the $J$ subintervals containing the signal; and take $I_j = [a_j,b_j]$. When we take the limit $n \to \infty$, we'll let $J = J_n$ grow with $n$, as is natural. We will see that we require $J_n = \Omega(n)$ in order for the limits to not vanish (and obviously $J_n = O(n)$ too).

@@

We fix a value $\Delta$ between $0$ and $L-1$. We will compute the second moment:
%
\begin{align}
%
    M_2(\Delta) = 
        \lim_{n \to \infty} \frac{1}{n} \sum_{i=1}^{n-\Delta} Y_i Y_{i+\Delta}.
%
\end{align}

\subsection{Clean signal without noise}

First, if there is no noise, and under the wide spacing assumption, we can break the sum into $J_n$ different sums, one for each copy of $x$ embedded in the signal:
%
\begin{align}
%
    M_2(\Delta) 
        &= \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=a_j}^{b_j-\Delta} Y_i Y_{i+\Delta}
        = \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
    \nonumber \\
    &= \frac{J_n L}{n} \frac{1}{L} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
        \to \gamma \cdot R_2(\Delta),
%
\end{align}
%
where here we have defined $R_2(\Delta)$ as element $\Delta$ of the autocorrelation of $x$, and $\gamma = \lim_{n\to\infty} J_n L / n$ is the fraction of the observations containing the signal.

\subsection{Pure noise without signal}

Here, $Y_i = \ep_i$. We first fix $\Delta$ between $1$ and $L-1$. Break up the sum into $\Delta$ terms as follows:
%
\begin{align}
%
    M_2(\Delta) &=\frac{1}{n} \sum_{i=1}^n \ep_i  \ep_{i+\Delta}
        \nonumber \\
    &= \frac{1}{\Delta} \sum_{m=0}^{\Delta-1} \frac{1}{n/\Delta}\sum_{j=1}^{n/\Delta} 
        \ep_{j\Delta + m} \ep_{(j + 1)\Delta + m}.
%
\end{align}
%
Each term
%
\begin{math}
    \frac{1}{n/\Delta}\sum_{j=1}^{n/\Delta} 
        \ep_{j\Delta + m} \ep_{(j + 1)\Delta + m}
\end{math}
%
is an average of $n / \Delta$ iid terms with expectation zero, and so converges to 0 a.s.\ as $n \to \infty$. (Of course, it's not exactly $n / \Delta$ terms; there will be finitely many terms unaccounted for this way, but these are negligible).

If $\Delta = 0$, then the computation is even easier:
%
\begin{align}
%
    M_2(\Delta) = \frac{1}{n} \sum_{i=1}^n \ep_i^2 
        \operatorname*{\longrightarrow}^{a.s.} \sigma^2.
%
\end{align}

So $M_2(\Delta) = \sigma^2$ if $\Delta = 0$, and 0 otherwise.

%%%So the second moment matrix of pure noise converges a.s.\ as $n \to \infty$ to $\sigma^2 I_L$.

\subsection{Signal plus noise}

This part is now easy. Denote by $\XX = x \ast g$, so $Y = \XX + \ep$. Then the second moment of the signal plus noise is:
%
\begin{align}
%
    M_2(\Delta,Y) = M_2(\Delta,\XX) + M_2(\Delta,\ep) 
        + \frac{1}{n}\sum_{i=1}^{n-\Delta} \XX_i \ep_{i+\Delta}
        + \frac{1}{n}\sum_{i=1}^{n-\Delta} \XX_{i+\Delta} \ep_i.
%
\end{align}
%
The law of large numbers says the cross terms vanish as $n \to \infty$. So the limit is simply:
%
\begin{align}
%
    M_2(\Delta,Y) = 
    \begin{cases}
        R_2(\Delta), &\text{ if } \Delta > 0; \\
        R_2(\Delta) + \sigma^2, &\text{ if } \Delta = 0.
    \end{cases}
%
\end{align}

\section{Third moments}

The same idea lets us compute the third moments. We fix two indices $\Delta_1 < \Delta_2$, and define:
%
\begin{align}
%
    M_3(\Delta_1,\Delta_2) = \sum_{i=1}^{n-\Delta_2} Y_i Y_{i+\Delta_1} Y_{i+\Delta_2}.
%
\end{align}

\section{Pure noise, no signal}

Again, the idea is to break up the big average into a fixed number of averages of $\approx n$ independent terms.

@@

### 2018-05-14 TEX/notes_will/notes_moments/backups/test2_save12.tex
updates on section 2

We have the following model. $x$ is a signal of length $L$. For large $n$, we observe the vector $Y \in \mathbb{R}^n$, which is of the form 
%
\begin{align}
%
    Y = x \ast g + N
%
\end{align}
%
where $g$ is a sum of diracs at locations at least $L$ apart, and $N$ is a vector of $n$ iid Gaussians. The question is, can we recover $x$?

We will assume that the first and last copies of $x$ is at least $L$ places removed from the boundaries of the interval $[1,n + L]$ (if not we can just zero-pad the interval). We will also denote by $I_1,\dots,I_J$ the $J$ subintervals containing the signal; and take $I_j = [a_j,b_j]$. When we take the limit $n \to \infty$, we'll let $J = J_n$ grow with $n$, as is natural. We will see that we require $J_n = \Omega(n)$ in order for the limits to not vanish (and obviously $J_n = O(n)$ too).

>>>>

We have the following model. $x$ is a signal of length $L$. For large $n$, we observe the vector $Y \in \mathbb{R}^n$, which is of the form 
%
\begin{align}
%
    Y = G \ast x + N
%
\end{align}
%
where $G$ is a sum of diracs at locations at least $L$ apart, and $N$ is a vector of $n$ iid Gaussians. The question is, can we recover $x$?

We will assume that the first and last copies of $x$ are at least $L$ places removed from the boundaries of the interval $[1,n + L]$ (if not we can just zero-pad the interval). We will also denote by $I_1,\dots,I_J$ the $J$ subintervals containing the signal; and take $I_j = [a_j,b_j]$. When we take the limit $n \to \infty$, we'll let $J = J_n$ grow with $n$, as is natural. We will see that we require $J_n = \Omega(n)$ in order for the limits to not vanish (and obviously $J_n = O(n)$ too).

### 2018-05-14 TEX/notes_will/notes_moments/backups/test2_save14.tex
updates on section 2

Again, the idea is to break up the big average into a fixed number of averages of $\approx n / \Delta_2$ independent terms. We can write:
%
\begin{align}
%
    & M_3(\Delta_1,\Delta_2) 
        \nonumber \\
    & = \frac{1}{\Delta_2}\sum_{m=0}^{\Delta_2-1}
        \frac{1}{(n/\Delta_2)}\sum_{j=1}^{n/\Delta_2} 
            N_{j+(j-1)\Delta_2+m} N_{j+\Delta_1 + (j-1)\Delta_2 + m} N_{j+j\Delta_2 + m}
%
\end{align}

>>>>

Again, the idea is to break up the big average over $n$ terms into a sum of $\Delta_2$ averages, each of $\approx n / \Delta_2$ independent terms. We write:
%
\begin{align}
%
    & M_3(\Delta_1,\Delta_2) 
        \nonumber \\
    & = \frac{1}{\Delta_2}\sum_{m=0}^{\Delta_2-1}
        \frac{1}{(n/\Delta_2)}\sum_{j=1}^{n/\Delta_2} 
            N_{j+(j-1)\Delta_2+m} N_{j+\Delta_1 + (j-1)\Delta_2 + m} N_{j+j\Delta_2 + m}.
%
\end{align}
%
Each of the $\Delta_2$ terms indexed by $m$ converges is an average of $n/\Delta_2$ independent terms with mean zero, and so converges a.s.\ to zero as $n \to \infty$.

### 2018-05-14 TEX/notes_will/notes_moments/backups/test2_save16.tex
updates on section 2

\newcommand{\F}{\mathcal{F}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\ep}{\varepsilon}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\GCD}{\mathbf{GCD}}
\newcommand{\XX}{\mathcal{X}}

@@

>>>>

\newcommand{\F}{\mathcal{F}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\ep}{\varepsilon}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\GCD}{\mathbf{GCD}}
\newcommand{\XX}{\mathcal{X}}

@@

We have:
%
\begin{align}
%
    M_3(\Delta_1,\Delta_2,Y) = M_3(\Delta_1,\Delta_2,\XX) + M_3(\Delta_1,\Delta_2,N)
        + \text{cross terms}.
%
\end{align}
%
If $0 < \Delta_1 < \Delta_2$, then all of the cross terms can be shown to go to zero by the same kind of argument we've used several times before.

If $\Delta_1 = \Delta_2 = \Delta$, there is a surviving cross-term, namely:
%
\begin{align}
%
    \frac{1}{n} \sum_{i=1}^{n-\Delta} \XX_i N_{i+\Delta}^2
    = \frac{1}{(n/L)} \sum_{j=1}^{J_n} \frac{1}{L} \sum_{i=a_j}^{b_j} \XX_i N_{i+\Delta}^2.
%
\end{align}
%
Each of $S_j \equiv \frac{1}{L} \sum_{i=a_j}^{b_j} \XX_i N_{i+\Delta}^2 \sim \frac{1}{L} \sum_{i=1}^{L} x_i \ep_i^2$, where $\ep_i \sim N(0,\sigma^2)$; and they are independent random variables with mean $\E[S_j] = \sigma^2 \cdot \bar{x}$. So by the law of large numbers,
%
\begin{align}
%
    \frac{1}{(n/L)} \sum_{j=1}^{J_n} \frac{1}{L} \sum_{i=a_j}^{b_j} \XX_i N_{i+\Delta}^2
    \to \gamma \cdot \sigma^2 \cdot \bar{x}.
%
\end{align}

Similarly, if $\Delta_1 = 0$, then there is a surviving cross-term which converges to $\gamma \cdot \sigma^2 \cdot \bar{x}$ as well. So in summary,
%
\begin{align}
%
    M_3(\Delta_1,\Delta_2,Y) = 
    \begin{cases}
        R_3(\Delta_1,\Delta_2), &\text{ if } 0 < \Delta_1 < \Delta_2; \\
        R_3(\Delta_1,\Delta_2) + \gamma \cdot \sigma^2 \cdot \bar{x},
                        &\text{ if } \Delta_1 = 0 \text{ or } \Delta_1 = \Delta_2.
    \end{cases}
%
\end{align}

%

### 2018-05-14 TEX/notes_will/notes_moments/backups/test2_save17.tex
updates on section 2

This part is now easy. Denote by $\XX = x \ast g$, so $Y = \XX + N$. Then the second moment of the signal plus noise is:
%
\begin{align}
%
    M_2(\Delta,Y) = M_2(\Delta,\XX) + M_2(\Delta,N) 
        + \frac{1}{n}\sum_{i=1}^{n-\Delta} \XX_i N_{i+\Delta}
        + \frac{1}{n}\sum_{i=1}^{n-\Delta} \XX_{i+\Delta} N_i.
%
\end{align}
%
The law of large numbers says the cross terms vanish as $n \to \infty$. So the limit is simply:
%
\begin{align}
%
    M_2(\Delta,Y) = 
    \begin{cases}
        R_2(\Delta), &\text{ if } \Delta > 0; \\
        R_2(\Delta) + \sigma^2, &\text{ if } \Delta = 0.
    \end{cases}
%
\end{align}

@@

Similarly, if $\Delta_1 = 0$, then there is a surviving cross-term which converges to $\gamma \cdot \sigma^2 \cdot \bar{x}$ as well. So in summary,
%
\begin{align}
%
    M_3(\Delta_1,\Delta_2,Y) = 
    \begin{cases}
        R_3(\Delta_1,\Delta_2), &\text{ if } 0 < \Delta_1 < \Delta_2; \\
        R_3(\Delta_1,\Delta_2) + \gamma \cdot \sigma^2 \cdot \bar{x},
                        &\text{ if } \Delta_1 = 0 \text{ or } \Delta_1 = \Delta_2.
    \end{cases}
%
\end{align}

@@

>>>>

This part is now easy. Denote by $\XX = x \ast g$, so $Y = \XX + N$. Then the second moment of the signal plus noise is:
%
\begin{align}
%
    M_2(\Delta,Y) = M_2(\Delta,\XX) + M_2(\Delta,N) 
        + \frac{1}{n}\sum_{i=1}^{n-\Delta} \XX_i N_{i+\Delta}
        + \frac{1}{n}\sum_{i=1}^{n-\Delta} \XX_{i+\Delta} N_i.
%
\end{align}
%
The law of large numbers says the cross terms vanish as $n \to \infty$. So the limit is simply:
%
\begin{align}
%
    M_2(\Delta,Y) = 
    \begin{cases}
        \gamma \cdot R_2(\Delta), &\text{ if } \Delta > 0; \\
        \gamma \cdot R_2(\Delta) + \sigma^2, &\text{ if } \Delta = 0.
    \end{cases}
%
\end{align}

@@

Similarly, if $\Delta_1 = 0$, then there is a surviving cross-term which converges to $\gamma \cdot \sigma^2 \cdot \bar{x}$ as well. So in summary,
%
\begin{align}
%
    M_3(\Delta_1,\Delta_2,Y) = 
    \begin{cases}
        \gamma \cdot R_3(\Delta_1,\Delta_2), &\text{ if } 0 < \Delta_1 < \Delta_2; \\
        \gamma \cdot R_3(\Delta_1,\Delta_2) + \gamma \cdot \sigma^2 \cdot \bar{x},
                        &\text{ if } \Delta_1 = 0 \text{ or } \Delta_1 = \Delta_2.
    \end{cases}
%
\end{align}

@@

\section{Attempt at EM}

Let's derive the EM algorithm, where we treat the locations $a_j$ as random latent variables. A simple model is that the $a_i$ are chosen independently and uniformly at random (though take note that this permits overlaps).

If $A = (a_1,\dots,a_J)$ is the vector of signal locations, then the generative model is:
%
\begin{align}
%
    p(Y , A| x) \propto 
%
\end{align}

### 2018-05-14 TEX/notes_will/notes_moments/backups/test2_save18.tex
updates on section 2

\newcommand{\F}{\mathcal{F}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\ep}{\varepsilon}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\GCD}{\mathbf{GCD}}
\newcommand{\XX}{\mathcal{X}}

@@

If $A = (a_1,\dots,a_J)$ is the vector of signal locations, then the generative model is:
%
\begin{align}
%
    p(Y , A| x) \propto 
%
\end{align}

>>>>

\newcommand{\F}{\mathcal{F}}
\newcommand{\LL}{\mathcal{L}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\ep}{\varepsilon}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\GCD}{\mathbf{GCD}}
\newcommand{\XX}{\mathcal{X}}

@@

We will denote by $G = (a_1,\dots,a_J)$ the random vector of signal locations; then the mean is $G \ast x $. We'll also assume the number $J$ of signals is known. The generative model is:
%
\begin{align}
%
    p(Y , G \, | \, x)
    = \frac{\exp\{-\| Y - G \ast x \|^2 / (2\sigma^2)\}}{(2\pi\sigma^2)^{n/2}} \frac{1}{n^J}.
%
\end{align}

The log-likelihood is then (up to an additive constant):
%
\begin{align}
%
    \LL(x; Y, G) \propto - \|Y - G\ast x\|^2.
%
\end{align}

### 2018-05-14 TEX/notes_will/notes_moments/backups/test2_save19.tex
updates on section 2

We have the following model. $x$ is a signal of length $L$. For large $n$, we observe the vector $Y \in \mathbb{R}^n$, which is of the form 
%
\begin{align}
%
    Y = G \ast x + N
%
\end{align}
%
where $G$ is a sum of diracs at locations at least $L$ apart, and $N$ is a vector of $n$ iid Gaussians. The question is, can we recover $x$?

@@

We'll compute the limiting vector of the sample average of all the windows. We fix an index $\ell$ between $1$ and $L$. The $\ell^{th}$ entry of the $i^{th}$ window is $Y_{i+\ell-1}$, $i=1,\dots,n$. Then the sample average of the $\ell^{th}$ entry is:
%
\begin{align}
%
    \frac{1}{n} \sum_{i=1}^n Y_{i+\ell - 1} &=
    \frac{1}{n}  \left( \sum_{i \notin I_j \forall j} N_i 
                  + \sum_{j=1}^J \sum_{k=1}^L (x[k] + N_{a_j + k-1} ) \right) 
    \nonumber \\
%
        &= \frac{1}{n}\sum_{i=1}^{n} N_i 
                  + \frac{J \cdot L}{n} \left(\frac{1}{L} \sum_{k=1}^L x[k]\right) 
    \nonumber \\
        &= \frac{1}{n}\sum_{i=1}^{n-L+1} N_i 
                  + \frac{J \cdot L}{n} \bar{x}       
     \nonumber \\
%
        &\operatorname*{\longrightarrow}^{a.s.}  \gamma  \cdot \bar{x}
%
\end{align}
%
where $\gamma = \lim_{n\to\infty} J_n L / n$; $\gamma$ is the fraction of the observations containing the signal. The limit of the $\ell^{th}$ entry of the mean does not depend on the index $\ell$, and is the same limit as if there were no noise at all, i.e.\ the noise completely averages out.

@@

First, if there is no noise, and under the wide spacing assumption, we can break the sum into $J_n$ different sums, one for each copy of $x$ embedded in the signal:
%
\begin{align}
%
    M_2(\Delta) 
        &= \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=a_j}^{b_j-\Delta} Y_i Y_{i+\Delta}
        = \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
    \nonumber \\
    &= \frac{J_n L}{n} \frac{1}{L} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
        \to \gamma \cdot R_2(\Delta),
%
\end{align}
%
where here we have defined $R_2(\Delta)$ as element $\Delta$ of the autocorrelation of $x$, and $\gamma = \lim_{n\to\infty} J_n L / n$ is the fraction of the observations containing the signal.

@@

This part is now easy. Denote by $\XX = x \ast g$, so $Y = \XX + N$. Then the second moment of the signal plus noise is:
%
\begin{align}
%
    M_2(\Delta,Y) = M_2(\Delta,\XX) + M_2(\Delta,N) 
        + \frac{1}{n}\sum_{i=1}^{n-\Delta} \XX_i N_{i+\Delta}
        + \frac{1}{n}\sum_{i=1}^{n-\Delta} \XX_{i+\Delta} N_i.
%
\end{align}
%
The law of large numbers says the cross terms vanish as $n \to \infty$. So the limit is simply:
%
\begin{align}
%
    M_2(\Delta,Y) = 
    \begin{cases}
        \gamma \cdot R_2(\Delta), &\text{ if } \Delta > 0; \\
        \gamma \cdot R_2(\Delta) + \sigma^2, &\text{ if } \Delta = 0.
    \end{cases}
%
\end{align}

@@

Let's derive the EM algorithm, where we treat the locations $a_j$ as random latent variables. A simple model is that the $a_i$ are chosen independently and uniformly at random (though take note that this permits overlaps).

We will denote by $G = (a_1,\dots,a_J)$ the random vector of signal locations; then the mean is $G \ast x $. We'll also assume the number $J$ of signals is known. The generative model is:
%
\begin{align}
%
    p(Y , G \, | \, x)
    = \frac{\exp\{-\| Y - G \ast x \|^2 / (2\sigma^2)\}}{(2\pi\sigma^2)^{n/2}} \frac{1}{n^J}.
%
\end{align}

@@

>>>>

We have the following model. $x$ is a signal of length $L$. For large $n$, we observe the vector $Y \in \mathbb{R}^n$, which is of the form 
%
\begin{align}
%
    Y = G \ast x + N
%
\end{align}
%
where $G$ is a sum of diracs at locations at least $2L$ apart, and $N$ is a vector of $n$ iid Gaussians. The goal is to recover $x$?

@@

We have:
%
\begin{align}
%
    M_1 = \frac{1}{n} \sum_{i=1}^n Y_i
        = \frac{1}{n/L} \sum_{j=1}^{J_n} \frac{1}{L}\sum_{i=1}^L x_i
            + \frac{1}{n} \sum_{i=1}^n N_i
    \to \gamma \cdot \bar{x},
%
\end{align}
%
where the limit is almost surely (we've used the strong law of large numbers), where $\gamma = \lim_{n\to\infty} J_n L / n$ is the fraction of the observations containing the signal.

@@

First, if there is no noise, and under the wide spacing assumption, we can break the sum into $J_n$ different sums, one for each copy of $x$ embedded in the signal:
%
\begin{align}
%
    M_2(\Delta) 
        &= \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=a_j}^{b_j-\Delta} Y_i Y_{i+\Delta}
        = \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
    \nonumber \\
    &= \frac{J_n L}{n} \frac{1}{L} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
        \to \gamma \cdot R_2(\Delta),
%
\end{align}
%
where here we have defined $R_2(\Delta)$ as element $\Delta$ of the autocorrelation of $x$. 
%
%
%

@@

We will denote by $\XX = x \ast G$, so $Y = \XX + N$. Then the second moment of the signal plus noise is:
%
\begin{align}
%
    M_2(\Delta,Y) = M_2(\Delta,\XX) + M_2(\Delta,N) 
        + \frac{1}{n}\sum_{i=1}^{n-\Delta} \XX_i N_{i+\Delta}
        + \frac{1}{n}\sum_{i=1}^{n-\Delta} \XX_{i+\Delta} N_i.
%
\end{align}
%
The law of large numbers says the cross terms vanish as $n \to \infty$. So the limit is simply:
%
\begin{align}
%
    M_2(\Delta,Y) = 
    \begin{cases}
        \gamma \cdot R_2(\Delta), &\text{ if } \Delta > 0; \\
        \gamma \cdot R_2(\Delta) + \sigma^2, &\text{ if } \Delta = 0.
    \end{cases}
%
\end{align}

@@

Let's derive the EM algorithm, where we treat the locations $a_j$ as random latent variables. A simple model is that the $a_j$ are chosen independently and uniformly at random (though take note that this permits overlaps of the signal copies).

We will denote by $G=\sum_{j=1}^J \delta_{a_j}$ the random vector of signal locations. We'll assume the number of signals $J$ is known. The generative model is:
%
\begin{align}
%
    p(Y , G | x)
    = \frac{\exp\{-\| Y - G \ast x \|^2 / (2\sigma^2)\}}{(2\pi\sigma^2)^{n/2}} 
        \cdot \frac{1}{n^J}.
%
\end{align}

@@

The probability distribution of $G$, given $Y$ and $x$, is:
%
\begin{align}
%
    p(G | Y, x) = \frac{ p(Y , G | x) }{ p(Y | x) }
    \propto \exp\{-\| Y - G \ast x \|^2 / (2\sigma^2)\}.
%
\end{align}

If $x^{(t)}$ is a guess for $x$ on the $t^{th}$ iterate of EM, then we define the weights:
%
\begin{align}
%
    w_t(G) = p(G | Y,x^{(t)}) = \exp\{-\| Y - G \ast x^{(t)} \|^2 / (2\sigma^2)\},
%
\end{align}
%
and the $Q$-function:
%
\begin{align}
%
    Q(x|x^{(t)}) 
    &= \E_{G|Y,x^{(t)}} \left[ \LL(x^{(t)};Y,G) \right]
        \nonumber \\
    &\propto - \sum_{G} \|Y - G\ast x\|^2 \cdot w_t(G).
%
\end{align}
%
The EM algorithm defines the next guess of $x$ to be:
%
\begin{align}
%
    x^{(t+1)} = \operatorname*{\arg\max}_x Q(x|x^{(t)})
        = \operatorname*{\arg\min}_x \sum_{G} \|Y - G\ast x\|^2 \cdot w_t(G).
%
\end{align}

But finding this minimum is hard, since the sum is over all ${n}\choose{J}$ possible values of $G$; if $J \sim n$, as we expect is necessary to have any chance of recovery, the number of terms will grow exponentially with $n$.

### 2018-05-14 TEX/notes_will/notes_moments/backups/test2_save20.tex
updates on section 2

%
%
%
%
\begin{document}

@@

We have the following model. $x$ is a signal of length $L$. For large $n$, we observe the vector $Y \in \mathbb{R}^n$, which is of the form 
%
\begin{align}
%
    Y = G \ast x + N
%
\end{align}
%
where $G$ is a sum of diracs at locations at least $2L$ apart, and $N$ is a vector of $n$ iid Gaussians. The goal is to recover $x$?

@@

Note that I'm not just computing the expected value. I'm actually showing convergence almost surely as $n \to \infty$.

@@

We will denote by $G=\sum_{j=1}^J \delta_{a_j}$ the random vector of signal locations. We'll assume the number of signals $J$ is known. The generative model is:
%
\begin{align}
%
    p(Y , G | x)
    = \frac{\exp\{-\| Y - G \ast x \|^2 / (2\sigma^2)\}}{(2\pi\sigma^2)^{n/2}} 
        \cdot \frac{1}{n^J}.
%
\end{align}

@@

But finding this minimum is hard, since the sum is over all ${n}\choose{J}$ possible values of $G$; if $J \sim n$, as we expect is necessary to have any chance of recovery, the number of terms will grow exponentially with $n$.

>>>>

%
%
%
%
\begin{document}
\title{Notes on big MRA}
\author{William Leeb}
\date{}
\maketitle

@@

We have the following model. $x$ is a signal of length $L$. For large $n$, we observe the vector $Y \in \mathbb{R}^n$, which is of the form 
%
\begin{align}
%
    Y = G \ast x + N
%
\end{align}
%
where $G$ is a sum of diracs at locations at least $2L$ apart, and $N$ is a vector of $n$ iid Gaussians. The goal is to recover $x$.

@@

For the moments computation, I'm not just computing the expected value. I'm actually showing convergence almost surely as $n \to \infty$. I probably made a mistake in the final formulas since I haven't check them carefully yet; the main point I want to record right now is how to get the bookkeeping correct for dealing with the noise terms, namely, we need to break the averages into averages of averages, each with iid terms.

@@

We will denote by $G=\sum_{j=1}^J \delta_{a_j}$ the random vector of signal locations. We'll assume the number of signals $J$ is known. The generative model is:
%
\begin{align}
%
    p(Y , G | x)
    = \frac{\exp\{-\| Y - G \ast x \|^2 / (2\sigma^2)\}}{(2\pi\sigma^2)^{n/2}} 
        \cdot {n \choose J}^{-1}.
%
\end{align}

@@

But finding this minimum is hard, since the sum is over all ${n}\choose{J}$ possible values of $G$; if $J \sim n$, as we expect is necessary to have any chance of recovery (this was suggested by the moments method), the number of terms will grow exponentially with $n$.

### 2018-05-14 TEX/notes_will/notes_moments/backups/test2_save3.tex
updates on section 2

First, if there is no noise, and under the wide spacing assumption, we can break the sum into $J_n$ different sums, one for each copy of $x$ embedded in the signal:
%
\begin{align}
%
    M_2(\Delta) 
        &= \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=a_j}^{b_j-\Delta} Y_i Y_{i+\Delta}
        = \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
    \nonumber \\
    &= \frac{J_n L}{n} \frac{1}{L} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
    \to \gamma \cdot m_2(\Delta),
%
\end{align}
%
where here we have defined $m_2(\Delta)$ as the $(1,1+\Delta)^{th}$ element of the second moment matrix of $\tilde{x} = [x,0] \in \mathbb{R}^{2L}$.

>>>>

First, if there is no noise, and under the wide spacing assumption, we can break the sum into $J_n$ different sums, one for each copy of $x$ embedded in the signal:
%
\begin{align}
%
    M_2(\Delta) 
        &= \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=a_j}^{b_j-\Delta} Y_i Y_{i+\Delta}
        = \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
    \nonumber \\
    &= \frac{J_n L}{n} \frac{1}{L} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
    \to \gamma \cdot m_2(\Delta),
%
\end{align}
%
where here we have defined $m_2(\Delta)$ as the $(1,1+\Delta)^{th}$ element of the second moment matrix $\tilde{x} \tilde{x}^\top / L$ of $\tilde{x} = [x,0] \in \mathbb{R}^{2L}$.

### 2018-05-14 TEX/notes_will/notes_moments/backups/test2_save4.tex
updates on section 2

First, if there is no noise, and under the wide spacing assumption, we can break the sum into $J_n$ different sums, one for each copy of $x$ embedded in the signal:
%
\begin{align}
%
    M_2(\Delta) 
        &= \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=a_j}^{b_j-\Delta} Y_i Y_{i+\Delta}
        = \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
    \nonumber \\
    &= \frac{J_n L}{n} \frac{1}{L} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
    \to \gamma \cdot m_2(\Delta),
%
\end{align}
%
where here we have defined $m_2(\Delta)$ as the $(1,1+\Delta)^{th}$ element of the second moment matrix $\tilde{x} \tilde{x}^\top / L$ of $\tilde{x} = [x,0] \in \mathbb{R}^{2L}$.

>>>>

First, if there is no noise, and under the wide spacing assumption, we can break the sum into $J_n$ different sums, one for each copy of $x$ embedded in the signal:
%
\begin{align}
%
    M_2(\Delta) 
        &= \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=a_j}^{b_j-\Delta} Y_i Y_{i+\Delta}
        = \frac{1}{n} \sum_{j=1}^{J_n} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
    \nonumber \\
    &= \frac{J_n L}{n} \frac{1}{L} \sum_{i=1}^{L-\Delta} x_i x_{i+\Delta}
        \to \gamma \cdot m_2(\Delta),
%
\end{align}
%
where here we have defined $m_2(\Delta)$ as element $\Delta$ of the (circular) autocorrelation of the zero-padded signal $\tilde{x} = [x,0] \in \mathbb{R}^{2L}$.
